### 2016-06-20 talk_Roscoff.tex
backslides

\documentclass[xcolor=table]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[british]{babel}
\usepackage[super]{nth}
%\usetheme{Boadilla}
%\usecolortheme{rose}
%\usecolortheme{crane}
%\usefonttheme{structuresmallcapsserif}
%\setbeamertemplate{navigation symbols}{}

\definecolor{Main}{rgb}{0.74, 0.13, 0.19}
\definecolor{Accent1}{rgb}{0.76,0.36,0.13}
\definecolor{Accent2}{rgb}{0.54,0.1,0.4}

\mode<presentation>{\usetheme{ilm}}
%\usecolortheme{rose}
%\useinnertheme[shadow]{circles}
%\usecolortheme{whale}
%\useoutertheme{infolines}

%\usecolortheme[named=Accent1]{structure}

%\setbeamerfont{page number in head/foot}{size=\large}
%\setbeamercolor{page number in head/foot}{fg=Main}
%% page/total
%%\setbeamertemplate{footline}[frame number]
%% pas de total
%\setbeamertemplate{footline}{%
%    	\hfill%
%	\usebeamercolor[fg]{page number in head/foot}%
%	\usebeamerfont{page number in head/foot}%
%	\insertframenumber\kern1em\vskip2pt%
%}
%
%\setbeamersize{text margin left=1em}
%\setbeamersize{text margin right=1em}

\usepackage[overlay,absolute]{textpos}
\setlength{\TPHorizModule}{10mm}
\setlength{\TPVertModule}{\TPHorizModule}
\textblockorigin{10mm}{10mm} % start everything near the top-left corner
\setlength{\parindent}{0pt}

%font
\usepackage[T1]{fontenc}
\usepackage{times}
%\usepackage[oldstylenums]{kpfonts}

%\include{macros}

%proper math and math symbols
%\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{datenumber,fp}

\usepackage[separate-uncertainty = true]{siunitx}

\usepackage{tabu}
\usepackage{multirow}
\usepackage{booktabs}

% Allow the usage of graphics (.jpg, .png, etc.) in the document
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, decorations.text, shapes.geometric, matrix,patterns,mindmap}
%\usetikzlibrary{positioning, patterns,topaths,chains,matrix}

\usepackage{pgfplots}
\usepackage{pgfplotstable}
\pgfplotsset{compat=1.9}
\usepgfplotslibrary{groupplots}
\usepgfplotslibrary{external}
\makeatletter
\newcommand*{\overlaynumber}{\number\beamer@slideinframe}
\tikzset{
  beamer externalizing/.style={%
    execute at end picture={%
      \tikzifexternalizing{%
        \ifbeamer@anotherslide
        \pgfexternalstorecommand{\string\global\string\beamer@anotherslidetrue}%
        \fi
      }{}%
    }%
  },
  external/optimize=false
}
\let\orig@tikzsetnextfilename=\tikzsetnextfilename
\renewcommand\tikzsetnextfilename[1]{\orig@tikzsetnextfilename{#1-\overlaynumber}}
\makeatother

\tikzset{every picture/.style={beamer externalizing}}
\tikzexternalize
\tikzsetexternalprefix{fig_presentation/}
%\tikzset{external/optimize=false}
%\tikzset{external/force remake}

%link or play movies
\usepackage{multimedia}

%beamer related package

\usepackage{todonotes}
\presetkeys{todonotes}{inline}{}

%bibliography
\usepackage[style=authoryear-comp, language=british,eprint=false, url=false, doi=false, sortcites=true, sorting=none, isbn=false, firstinits=true,maxcitenames=6]{biblatex}
%minimal citations
\AtEveryCitekey{%
	\clearfield{title}
	\clearfield{pages}
	\clearfield{volume}
	\clearfield{number}
	\clearfield{month}}
\newcommand{\myfullcite}[1]{{\scriptsize\fullcite{#1}}}
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{%
  \printtext{\bibstring{in}\intitlepunct}}}
%\bibliography{biblio}

\newcolumntype{P}[1]{>{\raggedright}p{#1}}

\institute[iLM]{iLM, Univ. Claude Bernard Lyon 1 and CNRS, France}
\title[crystal gel]{A novel route to the spontaneous formation of porous crystals via viscoelastic phase separation}
\author[M. Leocmach]{Mathieu Leocmach}
\date{30 June 2016}
\titlegraphic{
	\begin{tabu}{X[c]X[c]X[c]}
		\includegraphics[height=3\baselineskip]{presentation/Tsurusawa}&
		\includegraphics[height=3\baselineskip]{presentation/John}&
		\includegraphics[height=3\baselineskip]{presentation/Tanaka}\\
		Hideyo Tsurusawa & John Russo & Hajime Tanaka\\
		U. Tokyo & U. Bristol & U. Tokyo\\
	\end{tabu}

	%\vfill
	%\includegraphics[height=2\baselineskip]{logo_ens-lyon}\quad
	%\includegraphics[height=2\baselineskip]{logo_ums_grand}\quad
	%\includegraphics[height=2\baselineskip]{../../Yaourt/CNRSfilaire-Q}\quad
	%\includegraphics[height=2\baselineskip]{CRPP}\quad
	%\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{NEW-Logo-ERC-OUTLINE}
	}

\newlength{\mylength}

%\includeonly{creep_beamer}

\begin{document}
\tikzset{every mark/.append style={scale=0.8}}
\pgfplotsset{every axis/.append style={footnotesize}}

\pgfplotscreateplotcyclelist{earthy}{%
{red!40!black,mark=o},
{red!60!black,mark=triangle, every mark/.append style={rotate=180}},
{red!80!black,mark=square},
{red,mark=triangle},
{red!80!yellow, mark=diamond},
red!60!yellow,
red!40!yellow,
}

\AtBeginSection[]{
	\addtocounter{framenumber}{-1}
	\begin{frame}[plain]
		\tableofcontents[currentsection, hideothersubsections]
	\end{frame}
}

\begin{frame}{\pgfuseimage{cnrs-logo}\hspace*{0.3cm}\pgfuseimage{ucbl-logo}}%[plain]
	\titlepage
\end{frame}

\setcounter{framenumber}{0}

\begin{frame}{Crystallisation}
	\tikzsetnextfilename{whyx}%
	\input{presentation/whyx.pgf}

	\begin{block}{Nucleation}
	\begin{itemize}
		\item 10-1000 molecules $\rightarrow$ difficult to observe
		\item classical nucleation theory (CNT): fluid $\xrightarrow{\text{1 step}}$ crystal
	\end{itemize}
	\end{block}
\end{frame}

\begin{frame}{Let it snow!}
	\hfill Wegener-\alert{Bergeron}-Findeisen \alert{process}
	\begin{columns}[b]
	\column{0.4\textwidth}
	\includegraphics[width=\textwidth,clip=true,trim=0 0 0 1cm]{presentation/Pitter1.jpg}

	\vspace{1em}\includegraphics[width=\textwidth]{presentation/bergeron.jpg}
	\column{0.40\textwidth}
	\includegraphics[width=\textwidth,clip=true,trim=0 0 0.8mm 0]{presentation/queries-figure-1.jpg}

	\vspace{1em}\includegraphics[width=\textwidth]{presentation/OKC-Fallstreak.jpg}
	\end{columns}
\end{frame}

\begin{frame}{Colloid + Polymer}

\begin{columns}
	\column{0.45\textwidth}
	\tikzsetnextfilename{interactions}%
	\begin{tikzpicture}
		\let\mrad\relax
		\newlength\mrad
		\setlength{\mrad}{1em}

		%schemas interactions

		%repulsion
		\begin{scope}[xshift=2\mrad]
		\node[circle, inner sep=0, minimum width=4\mrad, inner color=Main, outer color=white]  (puri) {}; 
		\foreach \x in {0,30,...,330}{
			\draw[decoration={random steps,segment length=0.1\mrad, amplitude=0.08\mrad},decorate] (0,0) -- (\x:1\mrad);
		}
		\fill[gray, radius=0.8\mrad] circle;
		\foreach \x in {15,75,...,315}{
			\node[Accent2, font=\tiny] at (\x:0.6\mrad) {+};
		}
		\end{scope}

		%spheres collantes
		\begin{scope}[xshift=\textwidth-2.5\mrad]
		\node[draw,circle, inner sep=0, minimum width=2.5\mrad, green!50!black, dashed]  (colli) {}; 
		\draw[green!50!black, radius=1.25\mrad, dashed] (1.6\mrad,0) circle;
		\fill[gray, radius=0.8\mrad] circle;
		\fill[gray, radius=0.8\mrad] (1.6\mrad,0) circle;
		\end{scope}

		\path (puri) -- (colli) coordinate[midway] (m);

		%spheres dures
		\begin{scope}[shift=(m)]
		\node[circle, inner sep=0, minimum width=2.2\mrad, inner color=Main, outer color=white]  (duri) {}; 
		\foreach \x in {0,30,...,330}{
			\draw[decoration={random steps,segment length=0.1\mrad, amplitude=0.08\mrad},decorate] (0,0) -- (\x:1\mrad);
		}
		\fill[gray, radius=0.8\mrad] circle;
		\foreach \x in {15,75,...,315}{
			\node[Accent2, font=\tiny] at (\x:0.6\mrad) {+};
		}
		\end{scope}

		%fleches
		\draw[->] (puri) -- (duri) node[midway,above,font=\footnotesize]{ions};
		\draw[->] (duri) -- (colli)node[midway, above,font=\footnotesize]{PS};
		% \draw[->] (puri) -- (ps) -- (depli);

		%titres
		\node[text width=6em,anchor=base west, inner xsep=0] (pur) at (puri.west|-puri.north) {repulsion};

		\node[text width=6em,anchor=base, align=center] (dur) at (duri|-pur.base) {hard};

		\node[text width=6em, anchor=base east, inner xsep=0,xshift=1.5\mrad, align=right] (coll) at (colli.east|-pur.base) {depletion};
	\end{tikzpicture}\\

	\bigskip
	\includegraphics[width=\textwidth]{../CNRS-2015/Renth_sedim.jpg}\\
	\textit{\footnotesize Poon PRL 1999, Renth PRE 2001}
	\begin{itemize}
	\item triple coexistence
	\item various kinetic paths
	\end{itemize}

	\column{0.45\textwidth}
	\includegraphics[width=\textwidth,clip=true,trim=0 0 0 4.5cm]{presentation/Renth_diag.pdf}
	\end{columns}
\end{frame}

\begin{frame}{Gel formation}
	\begin{columns}
		\column{0.5\textwidth}
		\vspace{\baselineskip}
		\tikzsetnextfilename{phdiagth}%
		\begin{tikzpicture}[
			every axis/.style={
			xlabel absolute, every axis x label/.append style={anchor=base, yshift=-0.5em},
			ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
			}]
		\pgfplotscreateplotcyclelist{samples}{%
		blue,mark=square*,\\%
		cyan,mark=square\\%
		red, thick, mark=*\\%
		red!65!yellow, mark=otimes\\%
		red!30!yellow, mark=oplus\\%
		red!65!black, mark=o\\%
		}%
		\begin{groupplot}[
			group style={
					group name=g, group size=1 by 2,
					horizontal sep=4em,
					vertical sep=0.5em,
					x descriptions at=edge bottom,
				},
			cycle list name=samples,
			width=\textwidth,
			%height=0.5\linewidth,
			axis on top,
		]
		\nextgroupplot[
			ylabel={$\Pi v$},
			xmin=0,xmax=0.74,
			ymin=0, ymax=2400, %ytick={0,0.5,...,2},
			mark options={solid}
			]
			\addplot[black, dashed,no marks] file {res_gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
			\addplot[black, dashed,no marks] file {res_gasliquid_sl.phd};
			\addplot[black, no marks] file {res_gasliquid_bl.phd};
			\addplot[black, no marks] file {res_gasliquid_bg.phd};
			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd} node[above left, inner ysep=0, inner xsep=1pt] (x) at (rel axis cs:1,0) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
			\node at (axis cs:0.4,1500) {Gel};

		\nextgroupplot[
			xlabel={$\phi$}, ylabel={$c_p$ (\si{\milli\gram/\gram})},
			xmin=0,xmax=0.74,
			ymin=0, ymax=1.9, ytick={0,0.5,...,2},
			mark options={solid}
			]
			\addplot[black, dashed,no marks] file {gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
			\addplot[black, dashed,no marks] file {gasliquid_sl.phd};
			\addplot[black, no marks] file {gasliquid_bl.phd};
			\addplot[black, no marks] file {gasliquid_bg.phd};
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd} node[above left] (x) at (rel axis cs:1,0.5) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
			\draw[->] (x.base) --(rel axis cs:0.98,0.01);
		\end{groupplot}
		%\draw(g c1r1.outer south west) -- +(\textwidth,0);
		\end{tikzpicture}
		\column{0.5\textwidth}
		Small polymer/colloid size ratio
		\begin{itemize}
		\item metastable gas-liquid
		\item spinodal + glass $\Rightarrow$ gel\\ \hfill\textit{\footnotesize Lu et al. Nature~2008}
		\item continuous minority phase
		\item viscoelastic phase sep.\\ \hfill\textit{\footnotesize Tanaka J.Phys.Cond.Mat.~2000}
		\item due to dynamic assymetry
		\end{itemize}
		\tikzsetnextfilename{vps}%
		\begin{tikzpicture}
	\draw[->] (0,0) -- +(\textwidth,0) node[pos=0.05, label=-90:\textcolor{ilmorange}{$\tau_1$}](t1) {|} node[pos=0.54, label=-90:\textcolor{ilmcolor}{$\tau_2$}] (t2){|} coordinate[pos=1] (tf) node[anchor=north east,font=\footnotesize]{time};
	\let\side\relax
	\newlength\side
	\setlength{\side}{0.275\textwidth}
	\fill[ilmcolor!80!ilmorange] ($(t1)!0.1!45:(t2)$) coordinate(cse) rectangle +(\side,\side) coordinate(cnw);
	\begin{scope}
	\clip (cse) rectangle (cnw);
	\fill[ilmorange!80!ilmcolor] ++($(cse)!0.5!(cnw)$) 
		+(60:0.3\side) circle (0.22\side) 
		+(180:0.3\side) circle (0.22\side) 
		+(-60:0.3\side) circle (0.22\side) 
		+(0:0.6\side) circle (0.22\side)
		+(120:0.6\side) circle (0.22\side)
		+(-120:0.6\side) circle (0.22\side)
		+(45:0.8\side) circle (0.22\side)
		+(-45:0.8\side) circle (0.22\side);
	\end{scope}
	\fill[ilmcolor] (cse-|t2) ++(-0.5\side,0) coordinate(cse) rectangle +(\side,\side) coordinate(cnw);
	\begin{scope}
	\clip (cse) rectangle (cnw);
	\fill[ilmorange] ++($(cse)!0.5!(cnw)$) 
		+(60:0.3\side) circle (0.26\side) 
		+(180:0.3\side) circle (0.26\side) 
		+(-60:0.3\side) circle (0.25\side) 
		+(0:0.6\side) circle (0.25\side)
		+(120:0.6\side) circle (0.25\side)
		+(-120:0.6\side) circle (0.25\side)
		+(45:0.8\side) circle (0.25\side)
		+(-45:0.8\side) circle (0.25\side)
		%enhance broken bond
		+(120:0.15\side) circle (0.07\side);
	\end{scope}
	\fill[ilmorange] (cse-|tf) ++(-\side,0) coordinate(cse) rectangle +(\side,\side) coordinate(cnw);
	\begin{scope}
	\clip (cse) rectangle (cnw);
	\fill[ilmcolor] ++($(cse)!0.5!(cnw)$) circle (0.07\side)
		+(0:0.3\side) circle (0.07\side) 
		+(120:0.3\side) circle (0.07\side) 
		+(-120:0.3\side) circle (0.07\side) 
		+(-60:0.6\side) circle (0.07\side)
		+(60:0.6\side) circle (0.07\side)
		++(0:0.3\side)
		+(-60:0.3\side) circle (0.07\side)
		+(60:0.3\side) circle (0.07\side)
		+(120:0.6\side) circle (0.07\side)
		+(-120:0.6\side) circle (0.07\side)
		++(180:0.9\side)
		+(-60:0.3\side) circle (0.07\side)
		+(60:0.3\side) circle (0.07\side);
	\end{scope}
	\end{tikzpicture}

	\hfill$\textcolor{ilmcolor}{\tau_2} \rightarrow +\infty$ {\usebeamercolor[fg]{itemize item}\usebeamertemplate{itemize item}} arrest 
	\end{columns}
\end{frame}

\begin{frame}{Another class of gel? Crystal-gel?}
	\begin{itemize}
	\item Phase separation arrested by crystallisation
	\begin{description}
		\item[polymer blend] Tanaka \& Nishi \textit{PRL} 1985
		\item[simulations] Soga \textit{J.Chem. Phys.} 1999; Fortini \textit{PRE} 2008 
	\end{description}
	\item Spinodal then crystallisation
	\begin{description}
		\item[simulations] Pérez \textit{Langmuir} 2011
		\item[microgravity] (triple coexistence) Sabin \textit{PRL} 2012
	\end{description}
	\end{itemize}
	\includegraphics[width=\textwidth]{presentation/micrograv_Sabin2012}
\end{frame}

\begin{frame}{Experimental}
How to do that in 3D at particle level \alert{without a space station}?
\begin{itemize}
\item index matched + \alert{\SI{2.3}{\micro\metre} colloids} + confocal
\item density match + $T$-control
\item preparation protocol without flow
\end{itemize}
\bigskip
\begin{columns}
	\column{0.265\textwidth}
		Colloids+polymers\\
		{\footnotesize no screening
		\begin{itemize}
			\item Dispersed
			\item Homogeneous
		\end{itemize}
		\bigskip
		\begin{beamercolorbox}[rounded=true]{block body}PMMA + PS\\
		in decalin + CHB $\epsilon_r\approx5$\\
		$\kappa^{-1}>\SI{10}{\micro\metre}$
		\end{beamercolorbox}}
	\column{0.4\textwidth}
		\tikzsetnextfilename{chemin_gelification}%
		\input{presentation/chemin_gelification.pgf}
	\column{0.335\textwidth}
		\tikzsetnextfilename{cell}%
		\input{presentation/cell.pgf}
\end{columns}
\end{frame}

\begin{frame}{Exploring phase diagram}
	\vspace{\baselineskip}
	\tikzsetnextfilename{phdiag_cluster}%
	\begin{tikzpicture}[
		every axis/.style={
		xlabel absolute, every axis x label/.append style={anchor=base, yshift=-1em},
		%ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
		}]
	\pgfplotscreateplotcyclelist{samples}{%
	blue,mark=square*,\\%
	cyan,mark=square\\%
	red, thick, mark=*\\%
	red!65!yellow, mark=otimes\\%
	red!30!yellow, mark=oplus\\%
	red!65!black, mark=o\\%
	}%
	\begin{groupplot}[
		group style={
				group name=g, group size=2 by 2,
				horizontal sep=4em,
				vertical sep=3em,
				%x descriptions at=edge bottom,
			},
		cycle list name=samples,
		width=0.51\textwidth,
		height=0.5\linewidth,
		axis on top,
	]
	\nextgroupplot[
		title={phase diagram},
		xlabel={$\phi$}, ylabel={$c_p$ (\si{\milli\gram/\gram})},
		xmin=0,xmax=0.74,
		ymin=0, ymax=1.9, ytick={0,0.5,...,2},
		mark options={solid}
		]
		\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
		\addplot file {sample0.phd};
		\addplot file {sample1.phd};
		\addplot file {sample2.phd};
		\addplot file {sample3.phd};
		\addplot file {sample4.phd};
		\addplot file {sample5.phd};
		\end{scope}
		\addplot[black, dashed,no marks] file {gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
		\addplot[black, dashed,no marks] file {gasliquid_sl.phd};
		\addplot[black, no marks] file {gasliquid_bl.phd};
		\addplot[black, no marks] file {gasliquid_bg.phd};
		\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
		\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd} node[above left] (x) at (rel axis cs:1,0.5) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
		%\draw[gray] (axis cs:0.371290311343, 0.12299370925) -- (axis cs:0.680497991395,0.0372636873433);
		%\draw[gray, dotted] (axis cs:3.40152108798e-05,0.624739653238) -- (axis cs:0.722831684246, 0.0726528862772);
		\draw[->] (x.base) --(rel axis cs:0.98,0.01);
		\addplot[only marks, mark=triangle] file{fluid_samples.phd};

	\nextgroupplot[
		title={cluster size distribution},
		xlabel={$s$}, ylabel={$R_g/\sigma$},
		xmin=1, xmax=4e3,
		ymin=0.5, ymax=30,
		xmode=log, ymode=log, 
		clip mode=individual,
		]
		\addplot file{data/fig1b_sqrt_1.dat};
		\addplot file{data/fig1b_sqrt_2.dat};
		\addplot file{data/fig1b_sqrt_3.dat};
		\addplot file{data/fig1b_sqrt_4.dat};
		\addplot file{data/fig1b_sqrt_5.dat};
		\addplot[gray, ultra thick, domain=1:4e3] {0.629*x^(1/2.53)} node [left] at (rel axis cs:1,0.9){$D=2.53$};

	\end{groupplot}
	%\draw(g c1r1.outer south west) -- +(\textwidth,0);
	\end{tikzpicture}
	\begin{itemize}
	\item Good agreement with theoretical phase diagram
	\item At percolation time, compatible with random gelation
	\end{itemize}
\end{frame}

\begin{frame}{Outlier at later times}
	\vspace{\baselineskip}
	\tikzsetnextfilename{Sq_localphi}%
	\begin{tikzpicture}[
		every axis/.style={
		xlabel absolute, every axis x label/.append style={anchor=base, yshift=-1em},
		%ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
		}]
	\pgfplotscreateplotcyclelist{samples}{%
	blue,mark=square*,\\%
	cyan,mark=square\\%
	red, thick, mark=*\\%
	red!65!yellow, mark=otimes\\%
	red!30!yellow, mark=oplus\\%
	red!65!black, mark=o\\%
	}%
	\begin{groupplot}[
		group style={
				group name=g, group size=2 by 2,
				horizontal sep=4em,
				vertical sep=3em,
				%x descriptions at=edge bottom,
			},
		cycle list name=samples,
		width=0.51\textwidth,
		height=0.5\linewidth,
		axis on top,
	]
	\nextgroupplot[
		title={structure factor},
		xlabel={$q\sigma$}, ylabel={$S(q)$},
		xmode=log, ymode=log, xmin=0.2, xmax=20, ymin=0.1,
		restrict x to domain=-1.8:3, unbounded coords=discard, 
		%clip mode=individual
		]
		\addplot+[clip mode=individual] file{data/fig3c_1.dat};
		\addplot+[clip mode=individual] file{data/fig3c_2.dat};
		\addplot file{data/fig3c_3.dat};
		\addplot+[clip mode=individual] file{data/fig3c_4.dat};
		\addplot+[clip mode=individual] file{data/fig3c_5.dat};
		\addplot[black, ultra thick, domain=0.1:3] {2.14605*x^(-3)} node[pos=0.92,left=0.5em] {$D=3$};
		\addplot[gray, ultra thick, domain=0.1:3] {3.98797*x^(-2.53)} node [pos=0.65, above right] {$D=2.53$};

	\nextgroupplot[
		title={local volume fraction},
		xlabel={\# neighbours}, ylabel={$\phi$ [\%]},
		xmin=0, ymin=0, ytick={0,10,30,50},
		extra y ticks={64, 72}, extra y tick style={grid=major}, extra y tick label={},
		legend pos=north west, legend style={font=\footnotesize},
		no markers,
		/pgfplots/error bars/y dir=both, /pgfplots/error bars/y explicit,
		]
		\addplot table[y error=err]{data/vol_err_phi_0.10_cp_0.82.scaled};
		\addplot table[y error=err]{data/vol_err_phi_0.10_cp_1.36.scaled};
		\addplot table[y error=err]{data/vol_err_phi_0.25_cp_0.38.scaled};
		\addplot table[y error=err]{data/vol_err_phi_0.25_cp_0.48.scaled};
		\addplot table[y error=err]{data/vol_err_phi_0.25_cp_0.57.scaled};

	\end{groupplot}
	%\draw(g c1r1.outer south west) -- +(\textwidth,0);
	\end{tikzpicture}
	\begin{itemize}
	\item Final stucture compatible with random gelation
	\item Glassy local $\phi$ for 12-neighboured particles
	\end{itemize}
	\alert{Except} \textcolor{red}{$\bullet$}: compact structure and higher local $\phi$. %\textcolor{ilmorange}{$\Rightarrow$} Crystal?
\end{frame}

\begin{frame}{Real space: crystal gel!}
	\tikzsetnextfilename{phdiag_arrows}%
	\begin{tikzpicture}[
			every axis/.style={
			xlabel absolute, every axis x label/.append style={anchor=base, yshift=-1em},
			%ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
			}]
		\pgfplotscreateplotcyclelist{samples}{%
		blue,mark=square*,\\%
		cyan,mark=square\\%
		red, thick, mark=*\\%
		red!65!yellow, mark=otimes\\%
		red!30!yellow, mark=oplus\\%
		red!65!black, mark=o\\%
		}%
		\begin{axis}[
			cycle list name=samples,
			width=0.51\textwidth,
			height=0.5\linewidth,
			axis on top,
			xlabel={$\phi$}, ylabel={$c_p$ (\si{\milli\gram/\gram})},
			xmin=0,xmax=0.74,
			ymin=0, ymax=1.9, ytick={0,0.5,...,2},
			mark options={solid}
			]
			\begin{scope}[/pgfplots/skip coords between index={0}{1}, /pgfplots/skip coords between index={2}{3}]%[draw=none,mark repeat=2,mark phase=2]
			\addplot table {sample0.phd};
			\addplot table {sample1.phd};
			\addplot table {sample2.phd}[->] -- (rel axis cs:0.95,0.2);
			\addplot table {sample3.phd};
			\addplot table {sample4.phd};
			\addplot table {sample5.phd}[->] -- (rel axis cs:0.95,0.7);
			\end{scope}
			\addplot[black, dashed,no marks] file {gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
			\addplot[black, dashed,no marks] file {gasliquid_sl.phd};
			\addplot[black, no marks] file {gasliquid_bl.phd};
			\addplot[black, no marks] file {gasliquid_bg.phd};
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd} %node[above left] (x) at (rel axis cs:1,0.5) {Crystal}  
			-| (rel axis cs:1,0) \closedcycle;
			%\draw[->] (x.base) --(rel axis cs:0.98,0.01);
			\addplot[only marks, mark=triangle] file{fluid_samples.phd};
	\end{axis}
	\end{tikzpicture}
	\includegraphics[height=11\baselineskip,clip=true,trim=0.5cm 0 0 0]{nature/fig2.pdf}

	\bigskip
	Compact structure and higher local $\phi$ \alert{because crystallized}
\end{frame}

\begin{frame}{Crystals outgrow liquid}
\vspace{\baselineskip}
\tikzsetnextfilename{crystal_fates}%
\begin{tikzpicture}
\begin{groupplot}[%
			group style={
				group name=g, group size=2 by 1,
				horizontal sep=4em,
				},
		width=0.5\linewidth,
		%height=0.5\linewidth,
	]
	\nextgroupplot[
		title={nuclei},
		xlabel={$t/10^3\tau_B$}, ylabel=\#,
		xmin=0, xmax=13.0, ymin=0,
		no marks,
		legend style={cells={anchor=base west}},
		]
	\addplot table[x expr={\thisrowno{0}/2.3e3}]{data/fig3a_1.dat} node[left] at (rel axis cs:1,0.2) {avg size};
	\addplot table[x expr={\thisrowno{0}/2.3e3}]{data/fig3a_2.dat} node[left] at (rel axis cs:1,0.6) {number};
	%\legend{average size, average number};

	\nextgroupplot[
		title={cluster size distribution},
		xlabel={$s$}, ylabel=$R_g/\sigma$,
		xmode=log, ymode=log,
		ytick={0.5, 1, 2, 4}, yticklabels={$\frac{1}{2}$, 1, 2, 4},
		domain=1:5e2,
		]
	\addplot[only marks, red, thick, mark=*] file{data/fig3d_sqrt_1.dat};
	\addplot[black, ultra thick] {0.522*x^(1/3)} node[below left] at (axis cs:9e2,2.5) {$D=3$};
	\addplot[gray, ultra thick] {0.426*x^(1/2.53)} node[pos=0.08,below right] {$D=2.53$};
\end{groupplot}
\end{tikzpicture}
\begin{center}
\tikzsetnextfilename{pathsX}\vspace{-0.5em}\input{presentation/pathsX.pgf}
\end{center}
\end{frame}

\begin{frame}{Dynamics: Bergeron > Ostwald}
\vspace{\baselineskip}
\includegraphics[width=\textwidth]{nature/fig4.pdf}
\end{frame}

\begin{frame}{Conclusion}
	\begin{itemize}
		\item A particle level colloidal analogue to the Bergeron process
		\item Generic "crystal-gel" process
		\begin{enumerate}
			\item gas-liquid phase separation below melting point
		of a crystal
			\item weak or little frustration against crystallization
			\item dynamical slowing down in a supercooled liquid state
			\begin{itemize}
				\item viscoelastic phase separation
				\item network of the minority liquid phase
			\end{itemize}
			\item low supercooling: crystallisation > vitrification
		\end{enumerate}
		\item Gels of poor glass formers (monoatomic systems)
		\item Porous material with a large area of crystalline planes
		\begin{itemize}
			\item catalysis, sensing, filtration, super-capacitor\ldots
			\item single step
		\end{itemize}
	\end{itemize}
\end{frame}

\appendix
\newcounter{finalframe}
\setcounter{finalframe}{\value{framenumber}}

\begin{frame}[plain]
\end{frame}

\begin{frame}{Reproducibility in capillary}
	\hfill $\phi\approx 0.33$, $c_P=\SI{0.40}{\milli\gram/\gram}$

	\includegraphics[width=\textwidth]{nature/capillary}
\end{frame}

\begin{frame}{Transition probabilities}
	\vspace{\baselineskip}
	\includegraphics[width=\textwidth]{nature/sfig1.pdf}
\end{frame}

\begin{frame}{Evolution local volume fraction}
	\vspace{\baselineskip}
	12-neighboured particles
	\begin{center}
	\includegraphics{nature/local_phi12.pdf}
	\end{center}
\end{frame}

\begin{frame}{phase definition}
	\begin{description}
	\item[crystal]
		\begin{itemize} 
		\item particle $i$ has $N_b(i)=12$ neighbours
		\item $\displaystyle q_{6 m}(i)=\frac{1}{N_b(i)}\sum_{j=1}^{N_b(i)}
	Y_{6 m}(\mathbf{r_{ij}}),\quad -6 <m<6$
		\item crystalline bond $(i,j)$: \[(\mathbf{q}_{6}(i)/|\mathbf{q}_{6}(i)|)\cdot(\mathbf{q}_{6}(j)/|\mathbf{q}_{6}(j)|)>0.7\]
		\item crystalline particle: $7$ crystalline bonds
		\end{itemize}
	\item[surface] two neighbouring crystal particles
	\item[liquid] at least four neighbouring particles
	\item[gas]
	\end{description}
\end{frame}

\begin{frame}{Colloid and polymer sizes}
	Direct measurement \alert{too imprecise}
	\begin{itemize}
	\item diffusion constant $\Rightarrow R= \SI{2.3 +- 0.05}{\micro\metre}$
	\item Flory scaling litterature measurements $\Rightarrow R_g=\SI{76. +- 50.}{\nano\metre}$
	\end{itemize}

	\begin{block}{Adjusting theoretical and experimental phase diagrams}
	\begin{enumerate}
	\item to have the final local volume fraction of the 12-neighboured particles in the crystallising sample to be equal to the theoretical equilibrium crystal volume fraction 
	\item to have the theoretical spinodal line to correspond to the gel region boundary.
	\end{enumerate}
	\end{block}
	Starting from our initial estimates of $\sigma$ and $R$ we iteratively converge to $R=\SI{80}{\nano\metre}$, $\sigma = \SI{2.25}{\micro\metre}$ and a crystal (without defects) at $0.723$.
\end{frame}

\begin{frame}{Carnahan-Starling is not enough}
	\tikzsetnextfilename{CSphasediag}%
	\begin{tikzpicture}[
		every axis/.style={
			xlabel absolute, every axis x label/.append style={anchor=base, yshift=-0.5em},
			ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
			}]
		\pgfplotscreateplotcyclelist{samples}{%
		blue,mark=square*,\\%
		cyan,mark=square\\%
		red, thick, mark=*\\%
		red!65!yellow, mark=otimes\\%
		red!30!yellow, mark=oplus\\%
		red!65!black, mark=o\\%
		}%
		\begin{groupplot}[
			group style={
					group name=g, group size=2 by 2,
					horizontal sep=0.5em,
					vertical sep=0.5em,
					x descriptions at=edge bottom,
					y descriptions at=edge left,
				},
			cycle list name=samples,
			width=0.5\textwidth,
			%height=0.5\linewidth,
			axis on top,
			xlabel={$\phi$},
		]
		\nextgroupplot[
			ylabel={attraction},
			xmin=0,xmax=0.74,
			ymin=0, ymax=2400, %ytick={0,0.5,...,2},
			mark options={solid}
			]
			\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
		\addplot file {res_sample0.phd};
		\addplot file {res_sample1.phd};
		\addplot file {res_sample2.phd};
		\addplot file {res_sample3.phd};
		\addplot file {res_sample4.phd};
		\addplot file {res_sample5.phd};
		\end{scope}
			\addplot[black, dashed,no marks] file {res_gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
			\addplot[black, dashed,no marks] file {res_gasliquid_sl.phd};
			\addplot[black, no marks] file {res_gasliquid_bl.phd};
			\addplot[black, no marks] file {res_gasliquid_bg.phd};
			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} -- (rel axis cs:0,0) \closedcycle;
			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd}  -| (rel axis cs:1,0) \closedcycle;

	\nextgroupplot[
			xmin=0,xmax=0.74,
			ymin=0, ymax=2400, %ytick={0,0.5,...,2},
			mark options={solid}
			]
			\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
		\addplot file {res_sample0.phd};
		\addplot file {res_sample1.phd};
		\addplot file {res_sample2.phd};
		\addplot file {res_sample3.phd};
		\addplot file {res_sample4.phd};
		\addplot file {res_sample5.phd};
		\end{scope}
			\addplot[black, dashed,no marks] file {res_gasliquidCS_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
			\addplot[black, dashed,no marks] file {res_gasliquidCS_sl.phd};
			\addplot[black, no marks,skip coords between index={10}{100}] file {res_gasliquidCS_bl.phd};
			\addplot[black, no marks,skip coords between index={10}{100}] file {res_gasliquidCS_bg.phd};
			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} -- (rel axis cs:0,0) \closedcycle;
			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd}  -| (rel axis cs:1,0) \closedcycle;

		\nextgroupplot[
			ylabel={P concentr. (\si{\milli\gram/\gram})},
			xmin=0,xmax=0.74,
			ymin=0, ymax=1.9, ytick={0,0.5,...,2},
			mark options={solid}
			]
			\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
			\addplot file {sample0.phd};
			\addplot file {sample1.phd};
			\addplot file {sample2.phd};
			\addplot file {sample3.phd};
			\addplot file {sample4.phd};
			\addplot file {sample5.phd};
			\end{scope}
			\addplot[black, dashed,no marks] file {gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
			\addplot[black, dashed,no marks] file {gasliquid_sl.phd};
			\addplot[black, no marks] file {gasliquid_bl.phd};
			\addplot[black, no marks] file {gasliquid_bg.phd};
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} -- (rel axis cs:0,0) \closedcycle;
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd}  -| (rel axis cs:1,0) \closedcycle;
			\addplot[only marks, mark=triangle] file{fluid_samples.phd};

		\nextgroupplot[
			xmin=0,xmax=0.74,
			ymin=0, ymax=1.9, ytick={0,0.5,...,2},
			mark options={solid}
			]
			\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
			\addplot file {sample0.phd};
			\addplot file {sample1.phd};
			\addplot file {sample2.phd};
			\addplot file {sample3.phd};
			\addplot file {sample4.phd};
			\addplot file {sample5.phd};
			\end{scope}
			\addplot[black, dashed,no marks] file {gasliquidCS_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
			\addplot[black, dashed,no marks] file {gasliquidCS_sl.phd};
			\addplot[black, no marks,skip coords between index={10}{100}] file {gasliquidCS_bl.phd};
			\addplot[black, no marks,skip coords between index={10}{100}] file {gasliquidCS_bg.phd};
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} -- (rel axis cs:0,0) \closedcycle;
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd}  -| (rel axis cs:1,0) \closedcycle;
			\addplot[only marks, mark=triangle] file{fluid_samples.phd};
	\end{groupplot}
	\end{tikzpicture}
\end{frame}

\setcounter{framenumber}{\value{finalframe}}
\end{document}