### 2017-01-26 nature/RebuttalLetter.tex

\documentclass[11pt,a4paper]{article}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{xcolor}
\usepackage{setspace}

\usepackage[top=35mm,bottom=35mm,left=30mm,right=30mm]{geometry}

\newenvironment{referee}%
{\bigskip\singlespacing\bf}%
{\par\bigskip}

\begin{document}

%\doublespacing

%\doublespacing

\noindent
{\bf Dear Editor,}
\vskip 0.7cm
We would like to thank you for your providing us with two Referee reports for our manuscript (NM15113574). 
We are glad to see that both Referees 1 and 3 seem to have appreciated the improvements 
we made after the first round of reviews: Reviewer \#1 calls the work \emph{very beautiful and interesting}, while Referee \#3 goes into the details of explaining how we actually satisfactorily clarified her/his previous doubts, with just a couple of minor points left. Reviewer \#2 supported publication after some minor revisions in the first round of review.

After receiving the comments of the Referees' reports, we have taken them seriously and made an effort to address them properly. 
We summarize how we have addressed the critical comments of the Referees.
\begin{itemize}
 \item Referee \#1 seemed to think that the first part of the paper describes spinodal decomposition without showing anything new from previous studies (but she/he admits the novelty of the study of the different crystallzation channels). 
We are afraid that there may be some misunderstanding here. 
This is because the crystallisation of the liquid network requires an additional step, called stress-driven ageing, 
that was never observed experimentally before. However, this criticism makes us realize that this point was not so clear 
in the previous manuscript. 
To improve this, we spent considerable time developing a new analysis, which shows how spinodal decomposition proceeds differently depending on the polymer concentration, and that the ageing of the network due to mechanical stress is a necessary condition for the subsequent crystallization process. We think that this has considerably increased the impact of our work, because in addition to different crystallization channels, this additional work unveils a new link between the ageing of the liquid network and the subsequent crystallization process.
\item Referee \#3 suggests that we better acknowledge that the crystal-gel is found only in a quite specific region of the phase diagram. 
It's certainly true that crystal-gel states don't appear everywhere in the phase diagram, but this doesn't mean that they are hard to obtain. 
The Referee's comment led us to consider this problem in more depth. 
First we already showed that there is a reproducible region around the critical point were these states can be consistently obtained, and we were able to reproduce these findings both in our new cell design, and in ordinary capillaries, showing that the region of crystal-gel formation is well within the experimental range. Moreover, thanks to the above-mentioned new analysis, we have successfully unveiled the physical requirements for crystal-gel observation. This mechanism suggests that the extent of this region can be widened by changing the location of the critical point, which in our system is controlled by the size of the non-adsorbing polymer: moving the critical point to lower polymer concentrations opens the window where bonds can rearrange before the intervening glass transition. So we may say that the region of formation of crystal-gel porous structures is not only easily accessible, but also controllable.
\end{itemize}

We hope that the above explanation have addressed the comments of the Referees satisfactorily. We also took advantage of the new suggestions to improve the manuscript even further. In particular, as mentioned above, we have introduced a new type of analysis, which looks at the bond-breaking probability as a function of time, during the whole process of spinodal decomposition. Our new results show that, at low polymer-concentrations, the network undergoes a structural reorganisation, which results in more compact structures, that are a necessary condition for the appearance of the first crystalline nuclei. To our knowledge this is the first experimental demonstration of this process: it is also not easy to be captured by computer simulations, as they often neglect hydrodynamic interactions which play a fundamental role in this kinetic process, and no previous experiment has had the capability of tracking particles during the early stages of spinodal decomposition. The absence of any harmful flow upon the initiation of phase separation and any drift can be clearly seen in Supplementary Movie 1.

\vskip 0.3cm

Since the manuscript now includes the new analysis mentioned above and is expanded with 6 figures, we would like to resubmit it as an Article rather than a Letter. 
\textit{
\vskip 0.3cm}

We hope that you and your reviewers will find that the revised manuscript is now suitable for publication in Nature Materials.

\vskip 0.5cm

Warm thanks to you for your kind consideration and cooperation.

\vskip 0.8cm

\noindent
{\bf Yours Sincerely,
\vskip 0.3cm
\indent Hideyo Tsurusawa\\
\indent John Russo\\
\indent Mathieu Leocmach\\
\indent Hajime Tanaka}

\clearpage

\noindent
\begin{Large}
{\bf Response to Referee \#1}
\end{Large}

\vskip 0.5cm

We would like to thank the Referee for his/her comments. 
We are glad to see that he/she called the work \emph{very beautiful and interesting}. 
We address the concerns of the Referee in the following response.

\begin{referee}
The authors have improved their manuscript according to the comments of all referees, which has increased the readability of the paper. The novelty of the paper is also much better explained, which is in fact described not earlier than on page 7 of the manuscript.
The first 7 pages of the paper describes one of the arrest mechanisms of spinodal decomposition due to crystallization, which was already observed in experiments and simulations.
\end{referee}

We thank the Referee for his/her appreciation of the improvements made during our last revision of the manuscript. The Referee comments on the first 7 pages being not original, and simply describing the process of spinodal decomposition as previously observed in experiments and simulations. But we believe that our study offers novel insights also in the description of the early stages of the phase separation process. This is a very important point, as there is a clear link between the early-stage kinetics of the liquid network, and the late-stage crystallization process. In order to highlight this connection, in this new version of the manuscript we introduce a new type of analysis. In summary, we show that our work represents the first experimental demonstration of a process called \emph{stress-driven aging}. Most past simulation works have neglected, or not fully taken into account, the role of hydrodynamic interactions (HI) in the phase separation process. As was shown for example in Ref.~\cite{
furukawa2010key}, the phase separation process without HI produces rather compact structures, which would then easily crystallize to form compact nuclei. This is not what happens if one takes HI into consideration. Due to the incompressibility condition, the liquid network has a lower-connectivity, and the viscoelastic phase separation results in the accumulation of mechanical stress on these liquid bonds. Two things can happen depending on the polymer concentration. At high-polymer concentration, the bonds are not-easily broken, and the network which is first formed cannot undergo structural reorganisation. This inevitably leads to gelation. On the other side, if the polymer concentration is low enough, the liquid network can reorganise with a rather sharp transition to a more compact network. This condition is necessary for the appearance of the first crystalline nuclei. In our new manuscript, we are able to follow the kinetics of these processes by looking at the time evolution of the bond-breaking 
probability. The different network states are in fact characterised by different bond-breaking probabilities and by different coordination numbers. In Fig.~2a,b we show these quantities for high-polymer concentration samples, where spinodal decomposition is the only process that occurs. At low-polymer concentrations (Fig.~2c,d) instead, the network undergoes a sharp transition to a state with lower bond-breaking probability and higher connectivity. This transition is the stress-driven aging process, and increases the connectivity of the network. Despite being theoretically predicted, Ref.~\cite{tanaka2000,tanaka2007spontaneous}, this transition was not directly observed before.
We agree that the previous version of the manuscript was lacking on this part, so the Referee comments inspired us to develop the new analysis presented in new Fig. 2.

\begin{referee}
What is new is that crystallization not only takes place in the spinodal liquid branches, but can also occur via the Bergeron process in which the liquid evaporates, and the gas phase crystallizes again or via Ostwald ripening, i.e., melting of crystallites into the gas phase and recrystallization. Both processes make sense and can be explained that the crystallites should be surrounded by the gas phase due to the non-existence of a liquid-crystal coexistence as already speculated by Renth et al. It would be worthwhile to mention this more explicitly in the paper. In addition, the authors mention on page 3, that for confocal microscopy studies on colloidal gelation, the colloid-polymer mixture have so far been used. There are also confocal microscopy experiments performed on oppositely charged colloids,
which is worthwhile mentioning as well, Sanz et al., J. Phys. condens. Matt. 20, 494247 (2008), Sanz et al., J. Phys. Chem. B 112, 10861 (2008), where the effect of range and attraction strength have been studied on the different arrest mechanisms of the gels, including the crystallization of spinodal structures. Although the work is very beautiful and interesting, the referee has still strong doubts if the work provides sufficient novel insights to warrant publication in Nature Materials. First of all, the title suggests that a completely new route is found to form porous crystals. Crystallisation of spinodal networks has already been found in previous works and so the present work doesn't provide a new route, only another crystallization mechanism when the crystals get larger than the thickness of the branches. Secondly, the abstract suggests a new confocal microscopy protocol to study phase separation, which is in fact based on a salt injection that changes the interactions in
situ. Finally, the last sentence of the abstract is also misleading as this new route is not a single-step pathway to form nanoporous crystals, it involves spinodal decomposition, crystallization of the branches, and subsequently evaporation of the liquid and crystallization or melting of crystalline parts and recrystallization.
\end{referee}

In the new version of the manuscript we have tried to incorporate all the Referee's previous suggestions. 
Following the suggestion, we have explicitly mentioned the work of Poon and coworkers in connection to the Bergeron and Ostwald ripening processes.  
We have also added the two recommended references by Sanz et al. to the new version of the manuscript. 
On the doubt of the referee concerning the novelty of our work, we hope that our new analysis explained above provides a 
new physical insight on the physical condition required for the formation of porous crystal gels besides its new kinetic route. 
We also simplified the title to avoid misunderstanding of our message.

Regarding our experimental protocol, it is indeed reminiscent of salt injection, but differs in two important ways. (i) Rather than an injection salts are introduced through a membrane. This dialysis protocol allows salt introduction without external flow. (ii) We are in a regime where, upon salt introduction, charges are screened throughout the sample almost instantaneously, that is to say in a few Brownian times. This is due to the short length the ions have to diffuse from the reservoir to the top of the sample and to the large diffusion constant ratio between the ions and our large colloidal particles.
We believe this to be the first time this cell design has been applied for confocal microscopy observation at a single-particle level and used to study the formation of porous crystals in gels, which is the key to access the early stage of viscoelastic phase separation at a single particle resolution with full particle tracking. 
For example, the new analysis shown in new Fig.~2 is only possible with this new experimental protocol and full particle tracking.  
The absence of any harmful flow upon the initiation of phase separation and any drift can be clearly seen in Supplementary Movie 1.  
Finally, the last sentence of the abstract has been modified, following the kind suggestion of the Referee.

\vskip 0.5cm

We hope that the Referee will find that the new analysis have revealed a new mechanism behind 
crystal-gel formation and increased the novelty of our manuscript considerably and 
the revised manuscript is now suitable for publication in Nature Materials.

\clearpage

\noindent
\begin{Large}
{\bf Response to Referee \#3}
\end{Large}

\vskip 0.5cm

We would like to thank the Referee for his/her comments. In the following we respond to the new points raised.

\begin{referee}
\noindent
1) I was aware of the different frame rates on the two movies but I agree that the time stamps are helpful. Figure S5 addresses my concern about reproducibility at the single state point, and the different geometry itself is helpful too. I think the authors should be up front in the manuscript that the mechanism discussed is found in a very specific window (from the figures and SI, I think it is $\phi\sim 0.33$ and $c_p~0.39$ (plus minus) 0.01 mg/g) by moving this point into the abstract.
\end{referee}

\textcolor{blue}{We are glad to hear that the reviewer found that the time stamps are helpful and also Fig. S5 addresses his/her concern.}     
Following the Referee's recommendation, we have put the following sentence in the abstract: ``We clarify the physical requirements necessary for this to happen.'' Although this does not explicitly specify the points we observed crystal gels, our new analysis (see below) 
has revealed the physical conditions required for observing crystal-gel formation. We have also discussed this issue in our revised manuscript with the new analysis shown in a new Fig. 2. We stress that the conditions for crystal-gel formation are not stringent and well within the experimental accessible states, as we previously demonstrated in our reproducibility tests inside a capillary. Our new results show that, at low polymer-concentrations, the network undergoes a structural reorganisation, which results in more compact structures, that are a necessary condition for the appearance of the first crystalline nuclei. To our knowledge, this is the first experimental demonstration of this process: it is also not easy to be captured by computer simulations, as they often neglect hydrodynamic interactions which play a fundamental role in this kinetic process, and no previous experiment has had the capability of tracking particles during the early stages of spinodal decomposition. The absence of any harmful flow upon the initiation of phase separation and any drift can be clearly seen in Supplementary Movie 1.

\begin{referee}
2) The addition of the SI movie 3 and the discussion of Figure 2b addresses this concern for me. It's the percolation that is critical to the authors claim of a crystal gel - the percolated structure needs to persist over long times, otherwise, for example, the gel wouldn't have a finite low frequency elastic modulus. I think SI move 3 does show this.
\end{referee}

\textcolor{blue}{We are glad to hear that the reviewer thought that SI movie 3 and the discussion of Fig. 2b have addressed his/her concern and supported our claim of a crystal gel.}

\begin{referee}
3-1) The data quality of Figure 1c, Figure 1d appear insufficient to classify the $c_p$ = 0.38 mg/mg state point as different from the others. The authors should add replications, errors, and significance testing if they wish to distinguish the dimensionality (Fig 1c) and maximum volume fraction (Fig 1d) of the 0.38 mg/ml state point from the others.

I continue to think that the evidence for different behavior in the $c_p$ = 0.38 mg/g state point from Figure 1 c,d is weak. Supplemental Figure 3 is more convincing that the final state points are different. 
\end{referee}

We fully agree that the evidence from Fig.~1 c,d was not conclusive, but the goal of those figures was not to provide conclusive evidence, but to provide a comprehensive study of the early-stages of spinodal decomposition. In this context the Fig.~1 c,d only provide hints that the $c_p$ = 0.38 mg/g state point behaves differently from other state points. Direct evidence comes from the use of bond-orientational order parameters, presented in Fig. 5 and 6. 
Following the Referee's kind suggestion, we have now moved previous Fig. S3 to a new Fig. 3b. 
We hope that this will make the difference of the $c_p$ = 0.38 mg/mg state point from the others clearer.

\begin{referee}
3-2) (Yet, in the new section, the authors write: "This result clearly indicates that a single state point follows a different arrest mechanism, in which phase separation is arrested by crystallization and not by glassiness." Is that really true? It seems overstated. Don't Figure 1 c,d and SF3 just establish that the end point is different? They don't say anything about the mechanism, which would require kinetic analysis, as comes later. And it doesn't say anything about arrest, which would require dynamics.)
\end{referee}

In the new version of the manuscript we have removed that sentence, and replaced it with:

\emph{
These structural analysis provide some evidence that the state point $\phi\approx 0.33$ and $c_p=0.38\,$mg/g could be following a different arrest mechanism, in which phase separation is arrested by crystallization and not by glassiness. In the reminder of this manuscript we investigate this state more closely, to confirm these early results.}

We also agree that our previous manuscript didn't say much about the actual kinetic mechanism by which crystal-gels form. This point was also raised by Referee \#1 in noting that we didn't provide any additional insight in the phase separation process. In order to fully address this issue, we have implemented a kinetic analysis of the trajectories at all state points, computing the probability of bond-breaking between consecutive frames. This new kinetic analysis reveals that at low-polymer concentrations a new mechanism (stress-driven ageing) is taking place, which has the effect of increasing the average coordination of the colloidal particles in the liquid phase.
We believe this to be a novel observation of this phenomenon. Most past simulation works have neglected, or not fully taken into account, the role of hydrodynamic interactions (HI) in the phase separation process. As was shows for example in Ref.~\cite{furukawa2010key}, the phase separation process without HI produces rather compact structures, which would then easily crystallize to form compact nuclei. This is not what happens if one take HI into consideration. Due to the incompressibility condition, the liquid network has a lower-connectivity, and the viscoelastic phase separation results in the accumulation of mechanical stress on these liquid bonds. Two things can happen depending on the polymer concentration. At high-polymer concentration, the bonds are not-easily broken, and the network which is first formed cannot undergo structural reorganisation. This inevitably leads to gelation. On the other side, if the polymer concentration is low enough, the liquid network can reorganise with a rather sharp 
transition to a more compact network. This condition is necessary for the appearance of the first crystalline nuclei. In our new manuscript, we are able to follow the kinetics of these processes by looking at the time evolution of the bond-breaking probability. The different network states are in fact characterised by different bond-breaking probabilities and by different coordination numbers. In Fig.~2a,b we show these quantities for high-polymer concentration samples, where spinodal decomposition is the only process that occurs. At low-polymer concentrations (Fig.~2c,d) instead, the network undergoes a sharp transition to a state with lower bond-breaking probability and higher connectivity. This transition is the stress-driven aging process, and increases the connectivity of the network. Despite being theoretically predicted, Ref.~\cite{tanaka2000,tanaka2007spontaneous}, this transition was not directly observed before in experiments. 
We stress that for the experimental observation we need to follow particles at a single-particle level from the beginning of phase separation. This was made possible by a new experimental protocol of salt injection combined with confocal microscopy observation.

\begin{referee}
4) Figure S4 is critical to the interpretation, including information about all the transition pathways.
\end{referee}

\textcolor{blue}{We are glad to know that the reviewer thought that Fig. S4 is useful for supporting the interpretation.}

\begin{referee}
 5) The revised Figure 4, along with Figure S4, are very helpful to resolving the relative contributions of the three crystalline pathways. In Figure 4b,c, I still think that surface particles should somehow be distinguished - they represent a distinct state and a kind of consistency check, because the different crystallization pathways should all be mediated by this state. Without including the surface states in the primary figure, I don't see how the reader can inspect Figure S4, and get to the 54/34/11 abundances for direct, Bergeron and Ostwald processes. These abundances are the crucial claim of the paper.
 \end{referee}

 \textcolor{blue}{We are glad to hear that the revised Fig. 4, along with Fig. S4, are very useful to resolving the relative contributions of the three crystalline pathways. Following the kind suggestion of the reviewer, we have now included the fraction of surface particles in Fig. 6b and added the related discussion.}

\begin{referee}
In summary the confirmatory experiments (Fig S5 and Movie S3), the additional analysis of the crystalline state (Fig 3S) and the full transition analysis (rev Fig 4 and F S4) offer significant new support of the authors' claims; the authors have something new relative to prior literature. At this point, I think that the authors should: (i) better acknowledge that what is observed here is found in a quite specific region of the state diagram ($phi\sim 0.33$ and $c_p\sim 0.39$ {plus minus} 0.01 mg/g); (ii) better connect Figure 4 and Figure S4 by accounting for surface states in the former figure.

\end{referee}

We are glad to hear that the Referee thought that the previous revisions offered significant new support of the authors' claims. 
On the comment (i), as we explained above, we have clarified the physical conditions required for crystal-gel formation to take place. We have also discussed how we can widen the region where this is to be observed (see page 14). 
(ii)  \textcolor{blue}{As mentioned above,} following the suggestion, in the new version of the manuscript we follow the Referee's recommendation and include in Fig. 6b the fraction of surface particles, and update the description accordingly.

%\textcolor{red}{The Referee insists on having surface particles represented in Fig.6b. Even if I think that the figure becomes more complicated this way, maybe we should just update the figure. I'm sending Mathieu the dataset with the fraction of surface particles and liquid particles in case you agree on this change. If you do, I'll also update the description in the manuscript.}

%\textcolor{red}{We remind that the 54/34/11 abundances was obtained by looking at the phase history of each colloidal particle. Fig. 6b and Supplementary Fig. 4 don't contain information about the phase history of individual particles, just about the total number of particles in each phase, and the time-dependence of the transition probabilities. Of course these quantities are closely connected, and we hope to have provided a detailed description of the process (??).}

\vskip 5mm

We hope that the Referee will find that the new analysis has revealed a new mechanism behind 
crystal-gel formation and increased the novelty of our manuscript considerably and 
the revised manuscript is now suitable for publication in Nature Materials.

\clearpage

\bibliographystyle{naturemag2}
\bibliography{biblio}

\end{document}